Towards a free/open-source universal-dependency treebank for Kazakh

### 2015-06-21 2015-turklang-kazdep/paper.tex
small table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60915 72bbbca6-d526-0410-a7d9-f06f51895060

All of the traditionally defined ``cases'' in Kazakh have a variety of uses.  The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite (e.g., үйді \gloss{the house} \gmk{acc}, үйдің \gloss{of the house} \gmk{gen}).  This means that indefinite accusative and genitive as well as nominative noun phrases are all unmarked for case, and hence are all ambiguous (e.g., үй \gloss{house}).  The Kazakh transducer marks these all as \tag{nom}, but their use may be disambiguated.  Besides identifying the semantic role of an unmarked noun phrase, there are various agreement properties that make it clear what ``case'' the noun phrase ``truly'' receives: genitive noun phrases must have a corresponding noun phrase with possessive morphology that agrees in person with the genitive noun phrase; and truly nominative noun phrases must have a corresponding predicate that agrees in person with the nominative noun phrase---e.g., a verb or a copula.

>>>>

In Kazakh, as in most languages with case, there is not a one-to-one relation between ``case'' morphemes and syntactic function (not to mention a wide range of semantic functions).  The main syntactic functions of the traditionally defined cases in Kazakh are summarised in table~\ref{tab:cases}.

\begin{table}[htbp]
	\centering
	\caption{Primary syntactic functions of traditionally defined cases in Kazakh.}\label{tab:cases}
	\begin{small}
		\begin{tabular}{l p{5em} p{8em} p{22em}}
			\toprule
				\textbf{Case} & \textbf{Morpheme} & \textbf{Functions} & \textbf{Examples} \\
			\midrule
				nominative & - & subject,\newline
										attributive,\newline
										indefinite object,\newline
										indefinite genitival 
									& hargle bargle \newline
										jargal barble \newline 
										\\
				accusative & -/NI/ & object,\newline
											embedded subject,\newline
											causative subject
										& foo \newline
											bar \\
				genitive & -/NIң/ & genitival,\newline
											embedded subject \\
				locative & -/DA/ & adverbial
										& \\
				ablative & -/DAн/ & adverbial,\newline
											comparator \\
				dative & -/GA/ & indirect object,\newline
										attributive
									& \\
			\bottomrule
		\end{tabular}
	\end{small}
\end{table}

All of the traditionally defined ``cases'' in Kazakh have a variety of uses, .  The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite (e.g., үйді \gloss{the house} \gmk{acc}, үйдің \gloss{of the house} \gmk{gen}).  This means that indefinite accusative and genitive as well as nominative noun phrases are all unmarked for case, and hence are all ambiguous (e.g., үй \gloss{house}).  The Kazakh transducer marks these all as \tag{nom}, but their use may be disambiguated.  Besides identifying the semantic role of an unmarked noun phrase, there are various agreement properties that make it clear what ``case'' the noun phrase ``truly'' receives: genitive noun phrases must have a corresponding noun phrase with possessive morphology that agrees in person with the genitive noun phrase; and truly nominative noun phrases must have a corresponding predicate that agrees in person with the nominative noun phrase---e.g., a verb or a copula.
